---
title: "Reproducible Research Course Project 1"
author: "Robson"
date: "April 2, 2016"
---

###Loading and preprocessing the data

1. Loading the data
```{r}
activity <- read.csv("activity.csv")
```

2. Processing the data: converting the character variable "date" as Date format
```{r}
activity$date <- as.Date(as.character(activity$date))
```

###Total number of steps taken per day

1. Calculating the total number of steps taken per day
```{r}
stepsperday <- tapply(activity$steps, activity$date, sum, na.rm= TRUE)
```

2. Histogram of the total number of steps taken each day
```{r}
hist(stepsperday, col = "green", xlab = "Steps per Day", main = "Histogram of Steps per Day", ylim = c(0,30))
```

3. Calculating and reporting the mean and median of the total number of steps taken per day
```{r}
summary(stepsperday)
```

###Average daily activity pattern

1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
stepsaverage <- tapply(activity$steps, activity$interval, mean, na.rm= TRUE)
plot(names(stepsaverage), stepsaverage, type= "l", xlab = "Minutes past", ylab = "Steps Average taken each 5 min", main = "Steps taken averaged all days")
```

2. 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps:
```{r}
which.max(stepsaverage)
```

###Imputing missing values

1. Calculating and reporting the total number of missing values in the dataset
```{r}
sum(is.na(activity$steps))
```

2. Code to describe and show a strategy for imputing missing data. Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
activityimp <- activity
id <- dim(activityimp)[1]
for (i in 1:id){
        if (is.na(activityimp$steps[i])){
                int = activityimp$interval[i]
                activityimp$steps[i] = as.vector(stepsaverage[names(stepsaverage)==int])
        }
}
```

3. Making a histogram of the total number of steps taken each day and calculating and reporting the mean and median total number of steps taken per day.
```{r}
stepsperdayimp <- tapply(activityimp$steps, activityimp$date, sum, na.rm= TRUE)
hist(stepsperdayimp, col = "green", xlab = "Steps per Day", main = "Histogram of Steps per Day (with imputing missing data)")
summary(stepsperdayimp)
```
As we can see above, the histogram shape for imputed missing data looks like a normal curve compared with the histogram with missing data. The mean and median are a little bit bigger.

###Differences in activity patterns between weekdays and weekends

1. Creating a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r}
activityimp$daytype <- sapply(activityimp$date, function(x) if(weekdays(x) %in% c("sábado", "domingo")) x <- "weekend" else x <- "weekday")
activityimp$daytype <- as.factor(activityimp$daytype)
```

2. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
stepsaverageweekday <- tapply(activityimp$steps[activityimp$daytype=="weekday"], activityimp$interval[activityimp$daytype=="weekday"], mean, na.rm= TRUE)
stepsaverageweekend <- tapply(activityimp$steps[activityimp$daytype=="weekend"], activityimp$interval[activityimp$daytype=="weekend"], mean, na.rm= TRUE)
par(mfrow = c(2,1), cex = 0.7, pin = c(5.5,1.8))
plot(names(stepsaverageweekday), stepsaverageweekday, type= "l", ylab = "Steps Average", main = "Average of steps taken each 5 min: Weekday", ylim = c(0,230))
plot(names(stepsaverageweekend), stepsaverageweekend, type= "l", ylab = "Steps Average", main = "Weekend", ylim = c(0,230))
```

